Methodological Details - Blue Bananas

### 2020-10-30 RR_manuscript/Draft_RR.Rmd

bibliography      : ["r-references.bib"]

@@

# Introduction
In order to effectively accumulate knowledge, science needs to (i) produce data that can be recreated by using the same methods and (ii) arrive at conclusions about data that are robust. In recent coordinated efforts to replicate published findings, the scientific disciplines have uncovered surprisingly low succes rates for (i) (e.g., Open Science Collaboration 2015, Camerer et al. 2018, REF) leading to what is now referred to as the *replication crisis*. Beyond difficulties replicating scientific findings, more and more evidence suggests that the theoretical conclusions drawn from data are surprisingly variable even if researchers have access to reliable data (REFS). The latter situation has been referred to as the *inference crisis* (Rotello, Heit & Dubé 2015, Starns et al. 2019) and is, among other things, rooted in the inherent flexibility of data analysis, often referred to as researcher degrees of freedom (Simmons, Nelson, & Simonsohn, 2011, Gelman & Loken 2013). Data analysis involves many different steps including inspecting, organizing, transforming, and modeling the data. Along the way, different methodological and analytical choices need to be made, all of which might change its final interpretation. These researcher degrees of freedom are a blessing and a curse at the same time.

There are a blessing, because they allow us to look at nature from many different angles allowing us to make important discoveries and generate new hypothesis (e.g. Box 1976, Tukey 1977, de Groot 2014). They are a curse because idiosyncratic choices can lead to categorically different interpretation which find their way into the publication record where they are taken for granted (Simmons et al. 2011). Recent projects have shown that the variability between different data analysts is immense and leads researchers to draw vastly different conclusions about one and the same dataset (e.g. Silberzahn et al. 2018, Starns et al. 2019, Botvinik-Nezer et al., 2020). These projects, however, might still underestimate the extend to which analysts vary because data analysis is not only restricted to statistical inference of data tables. Human behavior is complex and offers many ways to be translated into numbers. This is particularly true for fields that draw conclusions about human behavior and cognition from multidimensional data like speech or video data. In fields working on human speech production for example, researchers need to make many decisions about what to measure and how to measure it. This is not trivial given the temporal extension of the acoustic signal and its complex structural composition. Decisions about operationalizing the raw data might not only influence downstream decisions about statistical modelling but statistical results might lead researchers to go back and revise earlier decisions about the raw data.

In this article, we investigate the diversity in analytic choices when many analyst teams analyze the same speech production data. We explore the interaction between analytic choices at the stage of the operationalization of raw data and subsequent statistical modelling. Specifically we report the impact of analytic pipeline on research results obtained by ## teams that gained access to the same raw data set to answer the same research question.

## Researcher degrees of freedom
Data analysis comes with many decisions like how to operationalize  a given phenomenon or behavior, what data to submit to statistical modelling and which to exclude, what models to use or what inferential decision procedure to apply.However, if these decisions during data analysis are not specified in advance, we might stumble upon seemingly meaningful patterns in the data that are merely statistical flukes. This can be problematic because to err is human.

@@

## Crowdsourcing alternative analyses
In a collaborative effort, Silberzahn et al. (2018) let twenty-nine independent analysis teams address the same research question. Analytic approaches and consequently the results varied widely between teams. 69% of the teams found support for the hypothesis, and 31% did not. Out of the 29 analysis strategies, there were 21 unique combinations of covariates. Importantly, the observed variability was neither predicted by the team's preconceptions about the phenomenon under investigation nor by peer ratings of the quality of their analyses. 
Their results suggest that analytic plurality is a fact of life and not driven by different levels of expertise or bias. Similar crowd-sourced studies recruiting independent analyst teams showed similar results.

@@

## Operationalizing speech 
*(copy pasted from RDF paper and slightly shortened by myself. I think its a good point of departure. Maybe one of your can try to rephrase it in your own words?)*

@@

## The data set - The prosody of redundant modifiers
Our data set was collected in order to answer the following research question: Do speakers acoustically modify a referring expression as a function of the typicality of the modifier of the noun (e.g. “a blue banana” vs. “a yellow banana”)?

@@

This account, however, has mainly focused on content selection (Gatt et al. 2013), i.e. whether a certain referential expression is chosen or not. However, speech communication is so much richer. Even looking at morphosyntactically identical expressions, speakers can modulate this expression via suprasegmental acoustic properties like temporal and spectral modifications of the segments involved (e.g. Ladd 2008). Most prominently, languages use intonation to signal discourse relationships between referents (among other functions). Intonation marks discourse-relevant referents for being new or given information to guide listeners’ interpretation of incoming messages. In many languages, speakers can use particular pitch movements to signal whether a referent has already been mentioned and is therefore referred back to, or a referent is newly introduced into the discourse. Many languages use intonation in order to signal if a referent is contrasting with one or more alternatives that are relevant to the current discourse. Content selection aside, in a scenario in which a speaker wants to refer to a banana when there is also a pear on the table, the speaker would most likely produce a high rising pitch accent on ‘banana’ to indicate the contrastive nature of the *noun.* In a scenario in which the speaker wants to refer to a yellow banana when there is also a less ripe green banana on the table, the speaker would most likely produce a high rising pitch accent on ‘yellow’ to indicate the contrastive nature of the *modifier.* In addition to a pitch accent, elements that are new and/or contrastive are often produced with additional suprasegmental prominence, i.e. segments are hyperarticulated, resulting in longer, louder and more clearly articulated acoustic targets.

@@

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

@@

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

>>>>

bibliography      : 
  - "r-references.bib"
  - "other_refs.bib"

@@

# Introduction

In order to effectively accumulate knowledge, science needs to (i) produce data that can be recreated by using the same methods and (ii) arrive at conclusions about data that are robust. 
In recent coordinated efforts to replicate published findings, the scientific disciplines have uncovered surprisingly low succes rates for (i) [e.g., @open2015estimating; @camerer2018evaluating] leading to what is now referred to as the *replication crisis*. 
Beyond difficulties replicating scientific findings, more and more evidence suggests that the theoretical conclusions drawn from data are surprisingly variable even if researchers have access to reliable data (REFS). 
The latter situation has been referred to as the *inference crisis* (Rotello, Heit & Dubé 2015, Starns et al. 2019) and is, among other things, rooted in the inherent flexibility of data analysis, often referred to as researcher degrees of freedom (Simmons, Nelson, & Simonsohn, 2011, Gelman & Loken 2013). 
Data analysis involves many different steps including inspecting, organizing, transforming, and modeling the data. 
Along the way, different methodological and analytical choices need to be made, all of which might change its final interpretation. 
These researcher degrees of freedom are a blessing and a curse at the same time.

There are a blessing, because they allow us to look at nature from many different angles allowing us to make important discoveries and generate new hypothesis (e.g. Box 1976, Tukey 1977, de Groot 2014). 
They are a curse because idiosyncratic choices can lead to categorically different interpretation which find their way into the publication record where they are taken for granted (Simmons et al. 2011). 
Recent projects have shown that the variability between different data analysts is immense and leads researchers to draw vastly different conclusions about one and the same dataset (e.g. Silberzahn et al. 2018, Starns et al. 2019, Botvinik-Nezer et al., 2020). 
These projects, however, might still underestimate the extent to which analysts vary because data analysis is not only restricted to statistical inference of data tables. 
Human behavior is complex and offers many ways to be translated into numbers. 
This is particularly true for fields that draw conclusions about human behavior and cognition from multidimensional data like speech or video data. 
In fields working on human speech production for example, researchers need to make many decisions about what to measure and how to measure it. 
This is not trivial given the temporal extension of the acoustic signal and its complex structural composition. 
Decisions about operationalizing the raw data might not only influence downstream decisions about statistical modelling but statistical results might lead researchers to go back and revise earlier decisions about the raw data.

In this article, we investigate the diversity in analytic choices when many analyst teams analyze the same speech production data. 
We explore the interaction between analytic choices at the stage of the operationalization of raw data and subsequent statistical modelling. 
Specifically we report the impact of analytic pipeline on research results obtained by XX teams that gained access to the same raw data set to answer the same research question.

## Researcher degrees of freedom

Data analysis comes with many decisions like how to operationalize a given phenomenon or behavior, what data to submit to statistical modelling and which to exclude, what models to use or what inferential decision procedure to apply. 
However, if these decisions during data analysis are not specified in advance, we might stumble upon seemingly meaningful patterns in the data that are merely statistical flukes. 
This can be problematic because to err is human.

@@

## Crowdsourcing alternative analyses

In a collaborative effort, Silberzahn et al. (2018) let twenty-nine independent analysis teams address the same research question. 
Analytic approaches and consequently the results varied widely between teams. 
Sixty-nine percent of the teams found support for the hypothesis, and 31% did not. 
Out of the 29 analysis strategies, there were 21 unique combinations of covariates. 
Importantly, the observed variability was neither predicted by the team's preconceptions about the phenomenon under investigation nor by peer ratings of the quality of their analyses. 
Their results suggest that analytic plurality is a fact of life and not driven by different levels of expertise or bias. 
Similar crowd-sourced studies recruiting independent analyst teams showed similar results.

@@

## Operationalizing speech

*(copy pasted from RDF paper and slightly shortened by myself. I think its a good point of departure. Maybe one of your can try to rephrase it in your own words?)*

@@

## The data set - The prosody of redundant modifiers

Our data set was collected in order to answer the following research question: Do speakers acoustically modify a referring expression as a function of the typicality of the modifier of the noun (e.g. “a blue banana” vs. “a yellow banana”)?

@@

This account, however, has mainly focused on content selection (Gatt et al. 2013), i.e. whether a certain referential expression is chosen or not. 
However, speech communication is so much richer. 
Even looking at morphosyntactically identical expressions, speakers can modulate this expression via suprasegmental acoustic properties like temporal and spectral modifications of the segments involved (e.g. Ladd 2008). 
Most prominently, languages use intonation to signal discourse relationships between referents (among other functions). 
Intonation marks discourse-relevant referents for being new or given information to guide listeners’ interpretation of incoming messages. 
In many languages, speakers can use particular pitch movements to signal whether a referent has already been mentioned and is therefore referred back to, or a referent is newly introduced into the discourse. 
Many languages use intonation in order to signal if a referent is contrasting with one or more alternatives that are relevant to the current discourse. 
Content selection aside, in a scenario in which a speaker wants to refer to a banana when there is also a pear on the table, the speaker would most likely produce a high rising pitch accent on ‘banana’ to indicate the contrastive nature of the *noun*. 
In a scenario in which the speaker wants to refer to a yellow banana when there is also a less ripe green banana on the table, the speaker would most likely produce a high rising pitch accent on ‘yellow’ to indicate the contrastive nature of the *modifier*. In addition to a pitch accent, elements that are new and/or contrastive are often produced with additional suprasegmental prominence, i.e. segments are hyperarticulated, resulting in longer, louder and more clearly articulated acoustic targets.

@@

# Methods

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

@@

## Data analysis

We used `r cite_r("r-references.bib")` for all our analyses.

### 2020-10-30 docs/ExperimentalDetails/Experimental_details.Rmd

## 1. Research Questions
This study investigates whether speakers modify the speech signal to signal atypical referents? Concretely, we ask whether the suprasegmental profile of an utterance with an atypical referent like "blue banana" is different from one with a typical referent such as "yellow banana".

## 2. Participants
30 native German speakers participated in this study. All participants grew up in a monolingual environment and were recruited from the population living in the broad Cologne area with normal or corrected-to-normal vision and normal hearing. Participants were paid for their participation.

## 3. Procedure
In this production study, participants interacted with the experimenter. Participants had to verbally instruct the experimenter to select a specified target object out of four visually presented objects. The non-target objects differed from the target with respect to their colour, their identity, or both. Objects were referred to using noun phrases that consist of a modifier denoting colour and a modified object (e.g. yellow banana, blue banana). These adjective-noun combinations differed with respect to the typicality of their combination. The adjective-noun combinations were either typical, medium typical, or atypical (as established in a norming study, see [additional materials](#adds)). Some combinations were typical such as yellow banana, red tomato, or yellow apricot; some were atypical such as red banana, purple tomato, or blue apricot; and some were somewhat in the middle like brown banana, green tomato, and red apricot.

@@

Participants were seated in front of a computer screen. The experimenter sat at the opposite side of the table in front of another computer screen. Participants and experimenter could see neither each other nor each others' screens. The experiment consisted of two phases. A familiarization phase and a test phase.

@@

“Du sollst den Würfel auf der grünen Sonnenbrille ablegen.”

*engl. 'You have to put the cube on top of the green sunglasses.'*

@@

“Und jetzt sollst du den Würfel auf der grünen Avokado ablegen.”

*engl. 'And now, you have to put the cube on top of the green avocado.'*

@@

The experimental program (implemented as a browser-based application) is available [here](https://github.com/SBRitter/fruits-production).

## Additional Information {#adds}
Here we specify in detail, how stimuli were selected for the above described production study.

### A.1 Selection of stimuli superset
The selection of stimuli was informed by a norming study for which we selected twenty fruits or vegetables (henceforth: FOODs) and four other referents (henceforth: NON-FOODs). Selection criteria were mainly informed by visual discriminability and phonemic composition, such that nouns denoting our objects were supposed to contain few voiceless segments.

@@

Blue: #29429f
Green: #3f8535
Red: #9d1c1c
Purple: #6d1b79
Brown: #2a1d11
Yellow: #e3c917
Orange: #ff8400
Orange (potatoes): #ff6600
Black: #000000
Grey: #FFFFFF

All pictures with natural and manipulated colors are available [here](https://osf.io/rdtx5/).

@@

### A.2 The norming study
100 German native speakers participated in a norming study using the crowd sourcing platform [Prolific]](https://www.prolific.ac). We presented all objects in different colors to our participants. FOODs came in 5 colors per object, NON-FOODs came in all nine colors. Participants were instructed to rate how typical they think the color for each object was, using a smooth slider ranging from 1 to 100 (see javascript for the experimental design [here]](https://github.com/stelaseldano/colour-typicality-norming). The results of the norming study can be retrieved [here](https://osf.io/znpg5/).

### A.3 Selection of experimental stimuli
The norming ratings were subsequently used to select appropriate stimuli for the production study according to the following procedure:

@@

atypical:
Yellow cherry (Gelbe Kirsche)
Green carrot (Grüne Möhre)
Red cucumber (Rote Gurke)
Orange grapes (Oangene Trauben)
Brown pepper (Braune Paprika)

medium:
Yellow peas (Gelbe Erbsen)
Green tomato (Grüne Tomate)
Red apricot (Rote Aprikose)
Orange potatoes (Orangene Kartoffeln)
Brown banana (Braune Banane)

typical:
Yellow lemon (Gelbe Zitrone)
Green green beans (Grüne Bohnen)
Red strawberry (Rote Erdbeere)
Orange mandarine (Orangene Mandarine)
Brown walnut (Braune Walnuss)

@@

Yellow sunglasses (Gelbe Sonnenbrille)
Yellow socks (Gelbe Socken)
Yellow clothes peg (Gelbe Wäscheklammer)
Green sunglasses (Grüne Sonnenbrille)
Green socks (Grüne Socken)
Green Paper clip (Grüne Büroklammer)
Red socks (Rote Socken)
Red paper clip (Rote Büroklammer)
Red clothes peg (Rote Wäscheklammer)
Orange socks (Orangene Socken)
Orange paper clip (Orangene Büroklammer)
Orange clothes peg (Orangene Wäscheklammer)
Brown sunglasses (Braune Sonnenbrille)
Brown paper clip (Braune Büroklammer)
Brown clothes peg (Braune Wäscheklammer)

@@

## A.4 Randomization
For each trial, targets, competitors, and distractors are chosen in a randomized way, fulfilling the following requirements:

@@

The python code for the randomization is available [here](https://github.com/SBRitter/fruits-randomization).

>>>>

## 1. Research Questions

This study investigates whether speakers modify the speech signal to signal atypical referents? 
Concretely, we ask whether the suprasegmental profile of an utterance with an atypical referent like "blue banana" is different from one with a typical referent such as "yellow banana".

## 2. Participants

Thirty native German speakers participated in this study. 
All participants grew up in a monolingual environment and were recruited from the population living in the broad Cologne area with normal or corrected-to-normal vision and normal hearing. 
Participants were paid for their participation.

## 3. Procedure

In this production study, participants interacted with the experimenter. 
Participants had to verbally instruct the experimenter to select a specified target object out of four visually presented objects. 
The non-target objects differed from the target with respect to their colour, their identity, or both. 
Objects were referred to using noun phrases that consist of a modifier denoting colour and a modified object (e.g. yellow banana, blue banana). 
These adjective-noun combinations differed with respect to the typicality of their combination. 
The adjective-noun combinations were either typical, medium typical, or atypical (as established in a norming study, see [additional materials](#adds)). 
Some combinations were typical such as yellow banana, red tomato, or yellow apricot; some were atypical such as red banana, purple tomato, or blue apricot; and some were somewhat in the middle like brown banana, green tomato, and red apricot.

@@

Participants were seated in front of a computer screen. 
The experimenter sat at the opposite side of the table in front of another computer screen. 
The participants and the experimenter could see neither each other nor each others' screens. 
The experiment consisted of two phases: a familiarization phase and a test phase.

@@

\noindent “Du sollst den Würfel auf der grünen Sonnenbrille ablegen.”  
\noindent *engl. 'You have to put the cube on top of the green sunglasses.'*

@@

\noindent “Und jetzt sollst du den Würfel auf der grünen Avokado ablegen.”  
\noindent *engl. 'And now, you have to put the cube on top of the green avocado.'*

@@

The experimental program (implemented as a browser-based application) is available here: [https://github.com/SBRitter/fruits-production](https://github.com/SBRitter/fruits-production).

## Additional Information {#adds}

Here we specify in detail, how stimuli were selected for the above described production study.

### A.1 Selection of stimuli superset

The selection of stimuli was informed by a norming study for which we selected twenty fruits or vegetables (henceforth: FOODs) and four other referents (henceforth: NON-FOODs). 
Selection criteria were mainly informed by visual discriminability and phonemic composition, such that nouns denoting our objects were supposed to contain few voiceless segments.

@@

\noindent Blue: #29429f  
\noindent Green: #3f8535  
\noindent Red: #9d1c1c  
\noindent Purple: #6d1b79  
\noindent Brown: #2a1d11  
\noindent Yellow: #e3c917  
\noindent Orange: #ff8400  
\noindent Orange (potatoes): #ff6600  
\noindent Black: #000000  
\noindent Grey: #FFFFFF

All pictures with natural and manipulated colors are available here: [https://osf.io/rdtx5/](https://osf.io/rdtx5/).

@@

### A.2 The norming study

One-hundred German native speakers participated in a norming study using the crowd sourcing platform [Prolific]](https://www.prolific.ac). 
We presented all objects in different colors to our participants. 
FOODs came in 5 colors per object, NON-FOODs came in all nine colors. 
Participants were instructed to rate how typical they think the color for each object was, using a smooth slider ranging from 1 to 100 (see javascript for the experimental design here: [https://github.com/stelaseldano/colour-typicality-norming](https://github.com/stelaseldano/colour-typicality-norming). 
The results of the norming study can be retrieved here: [https://osf.io/znpg5/](https://osf.io/znpg5/).

### A.3 Selection of experimental stimuli

The norming ratings were subsequently used to select appropriate stimuli for the production study according to the following procedure:

@@

\noindent atypical:  
\noindent Yellow cherry (Gelbe Kirsche)  
\noindent Green carrot (Grüne Möhre)  
\noindent Red cucumber (Rote Gurke)  
\noindent Orange grapes (Oangene Trauben)  
\noindent Brown pepper (Braune Paprika)

\noindent medium:  
\noindent Yellow peas (Gelbe Erbsen)  
\noindent Green tomato (Grüne Tomate)  
\noindent Red apricot (Rote Aprikose)  
\noindent Orange potatoes (Orangene Kartoffeln)  
\noindent Brown banana (Braune Banane)

\noindent typical:  
\noindent Yellow lemon (Gelbe Zitrone)  
\noindent Green green beans (Grüne Bohnen)  
\noindent Red strawberry (Rote Erdbeere)  
\noindent Orange mandarine (Orangene Mandarine)  
\noindent Brown walnut (Braune Walnuss)

@@

\noindent Yellow sunglasses (Gelbe Sonnenbrille)  
\noindent Yellow socks (Gelbe Socken)  
\noindent Yellow clothes peg (Gelbe Wäscheklammer)  
\noindent Green sunglasses (Grüne Sonnenbrille)  
\noindent Green socks (Grüne Socken)  
\noindent Green Paper clip (Grüne Büroklammer)  
\noindent Red socks (Rote Socken)  
\noindent Red paper clip (Rote Büroklammer)  
\noindent Red clothes peg (Rote Wäscheklammer)  
\noindent Orange socks (Orangene Socken)  
\noindent Orange paper clip (Orangene Büroklammer)  
\noindent Orange clothes peg (Orangene Wäscheklammer)  
\noindent Brown sunglasses (Braune Sonnenbrille)  
\noindent Brown paper clip (Braune Büroklammer)  
\noindent Brown clothes peg (Braune Wäscheklammer)

@@

## A.4 Randomization

For each trial, targets, competitors, and distractors are chosen in a randomized way, fulfilling the following requirements:

@@

The python code for the randomization is available here: [https://github.com/SBRitter/fruits-randomization](https://github.com/SBRitter/fruits-randomization).
